---
title: "InitialResearch"
author: "Trevor Isaacson"
date: "3/27/2022"
output: html_document
---


Machine Learning and Cosine Similarity? Ask about slides
Difference in measurement start of day and end of day, compare that number with snowfall two weeks out
Average measurement compared with snowfall two weeks out 
```{r setup, include=FALSE, echo = FALSE}
set.seed(445)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width=5, fig.height=3) 
knitr::opts_chunk$set(message = FALSE)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(e1071)
library(rmarkdown)
library(glmnet)
library(knitr)
library(dplyr)
library(zoo)
```

## Original Buoy Data
```{r}
originalBuoy2019 = read.delim("./buoyData/originalBuoy2019.txt", header = FALSE, sep = "")
colnames(originalBuoy2019) = c("Year", "Month", "Day", "Hour", "Minute", "WindDir", "WindSpeed", "GustSpeed", "WaveHeight", "DominantWavePeriod", "AverageWavePeriod", "DirWavePeriod", "SeaLevelPressure", "AirTemp", "SeaSurfaceTemp", "DewpointTemp", "Visibility", "Tide")
originalBuoy2019 = originalBuoy2019[-c(1),]

originalBuoy2020 = read.delim("./buoyData/originalBuoy2020.txt", header = FALSE, sep = "")
colnames(originalBuoy2020) = c("Year", "Month", "Day", "Hour", "Minute", "WindDir", "WindSpeed", "GustSpeed", "WaveHeight", "DominantWavePeriod", "AverageWavePeriod", "DirWavePeriod", "SeaLevelPressure", "AirTemp", "SeaSurfaceTemp", "DewpointTemp", "Visibility", "Tide")
originalBuoy2020 = originalBuoy2020[-c(1),]
```

```{r}
# the measurements we want are measured at minute 40
originalBuoy2019 = originalBuoy2019 %>%
  # these aren't measured and remain static
  select(-c("Tide", "Visibility")) %>%
  filter(Minute == 40)

originalBuoy2019 = originalBuoy2019 %>%
  mutate(WindSpeed = as.numeric(as.character(WindSpeed))) %>%
  mutate(GustSpeed = as.numeric(as.character(GustSpeed))) %>%
  mutate(AirTemp = as.numeric(as.character(AirTemp))) %>%
  mutate(SeaSurfaceTemp = as.numeric(as.character(SeaSurfaceTemp))) %>%
  mutate(DewpointTemp = as.numeric(as.character(DewpointTemp))) %>%
  mutate(WaveHeight = as.numeric(as.character(WaveHeight))) %>%
  mutate(DominantWavePeriod = as.numeric(as.character(DominantWavePeriod))) %>%
  mutate(AverageWavePeriod = as.numeric(as.character(AverageWavePeriod))) %>%
  mutate(DirWavePeriod = as.numeric(as.character(DirWavePeriod))) %>%
  mutate(SeaLevelPressure = as.numeric(as.character(SeaLevelPressure))) %>%
  mutate(WindDir = as.numeric(as.character(WindDir))) %>%
  mutate(Day = as.numeric(as.character(Day))) %>%
  mutate(Month = as.numeric(as.character(Month))) %>%
  mutate(Year = as.numeric(as.character(Year))) %>%
  filter(WaveHeight != 99.00)

originalBuoy2019


originalBuoy2020 = originalBuoy2020 %>%
  # these aren't measured and remain static
  select(-c("Tide", "Visibility")) %>%
  filter(Minute == 40)

originalBuoy2020 = originalBuoy2020 %>%
  mutate(WindSpeed = as.numeric(as.character(WindSpeed))) %>%
  mutate(GustSpeed = as.numeric(as.character(GustSpeed))) %>%
  mutate(AirTemp = as.numeric(as.character(AirTemp))) %>%
  mutate(SeaSurfaceTemp = as.numeric(as.character(SeaSurfaceTemp))) %>%
  mutate(DewpointTemp = as.numeric(as.character(DewpointTemp))) %>%
  mutate(WaveHeight = as.numeric(as.character(WaveHeight))) %>%
  mutate(DominantWavePeriod = as.numeric(as.character(DominantWavePeriod))) %>%
  mutate(AverageWavePeriod = as.numeric(as.character(AverageWavePeriod))) %>%
  mutate(DirWavePeriod = as.numeric(as.character(DirWavePeriod))) %>%
  mutate(SeaLevelPressure = as.numeric(as.character(SeaLevelPressure))) %>%
  mutate(WindDir = as.numeric(as.character(WindDir))) %>%
  mutate(Day = as.numeric(as.character(Day))) %>%
  mutate(Month = as.numeric(as.character(Month))) %>%
  mutate(Year = as.numeric(as.character(Year))) %>%
  filter(WaveHeight != 99.00)

originalBuoy2020
```

```{r}
# create mean for every 4 hours, 12 hours and day
DailyData2019 = originalBuoy2019 %>%
  group_by(Year, Month, Day) %>%
  summarise(across(c(WindDir:DewpointTemp), ~ mean(.x, na.rm = TRUE)))
  
DailyData2020 = originalBuoy2020 %>%
  group_by(Year, Month, Day) %>%
  summarise(across(c(WindDir:DewpointTemp), ~ mean(.x, na.rm = TRUE)))

DailyData2019
DailyData2020
```

## Park City Snowfall Data
```{r}
ParkCity2019 = read.csv("./snowfallData/ParkCitySnowfall2019.csv")
ParkCity2020 = read.csv("./snowfallData/ParkCitySnowfall2020.csv")
ParkCity2021 = read.csv("./snowfallData/ParkCitySnowfall2021.csv")
```

```{r warning = FALSE}
# change M, T and - to 0
ParkCity2019 = ParkCity2019 %>%
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.character, as.numeric)
ParkCity2019[is.na(ParkCity2019)] <- 0

ParkCity2020 = ParkCity2020 %>%
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.character, as.numeric)
ParkCity2020[is.na(ParkCity2020)] <- 0

ParkCity2019
ParkCity2020

```

The Dataframe ParkCity2019 shows snowfall totals for the corresponding day.  
Snowfall column shows the snowfall on that day.  
Snowfall14DaysLater shows the snowfall two weeks after that day.  This will be used to determine if waveHeight increases in Hawaii lead to snow 14 days later in Utah
```{r warning = FALSE}
# make snowtotals into one column
snowTotals2019 = stack(ParkCity2019[2:ncol(ParkCity2019)])
snowTotals2019 = snowTotals2019[-c(60,61,62,124,168,186,248,341), ]

# advanced snowfall = park city snow two weeks in advance
# remove first 14 days of year
advanceSnowTotals14Days = snowTotals2019[-c(1:14), ]
# add first 14 days of next year to end
SnowTotals14Daysinto2020 = ParkCity2020[1:14, "Jan"]
SnowTotals14Daysinto2020 = data.frame(values = SnowTotals14Daysinto2020, 
                                      ind = "Jan")
advancedSnowfall = rbind(advanceSnowTotals14Days, SnowTotals14Daysinto2020)

# create dataframe of snowfalls 
newParkCity2019 = data.frame(Year = DailyData$Year,
                             Month = DailyData$Month,
                             Day = DailyData$Day,
                             Snowfall = snowTotals2019$values,
                             Snowfall14DaysLater = advancedSnowfall$values)
ParkCity2019 = newParkCity2019

DailyData2019["Snowfall"] = ParkCity2019$Snowfall
DailyData2019["Snowfall14DaysLater"] = ParkCity2019$Snowfall14DaysLater

DailyData2019
```

```{r}
# make snowtotals into one column
snowTotals2020 = stack(ParkCity2020[2:ncol(ParkCity2020)])
snowTotals2020
snowTotals2020 = snowTotals2019[-c(60,61,124,168,186,248,341), ]

# advanced snowfall = park city snow two weeks in advance
# remove first 14 days of year
advanceSnowTotals14Days = snowTotals2020[-c(1:14), ]
# add first 14 days of next year to end
# SnowTotals14Daysinto2020 = ParkCity2020[1:14, "Jan"]
# SnowTotals14Daysinto2020 = data.frame(values = SnowTotals14Daysinto2020, ind = "Jan")
# advancedSnowfall = rbind(advanceSnowTotals14Days, SnowTotals14Daysinto2020)

# create dataframe of snowfalls 
# newParkCity2019 = data.frame(Year = DailyData$Year,
#                              Month = DailyData$Month,
#                              Day = DailyData$Day,
#                              Snowfall = snowTotals2019$values,
#                              Snowfall14DaysLater = advancedSnowfall$values)
# ParkCity2019 = newParkCity2019

# DailyData2019["Snowfall"] = ParkCity2019$Snowfall
# DailyData2019["Snowfall14DaysLater"] = ParkCity2019$Snowfall14DaysLater

# DailyData2019




```

```{r}
ggplot(data = DailyData2019, aes(x = Day, y = WaveHeight)) +
  geom_line() +
  facet_wrap(~Month)

ggplot(data = ParkCity2019, aes(x = Day, y = Snowfall14DaysLater)) +
  geom_line() +
  facet_wrap(~Month)
```


```{r}
predictedSnow = lm(Snowfall14DaysLater ~ SeaLevelPressure , data = DailyData2019)
summary(predictedSnow)


```










